---
title: "DailyCheck#9"
output: pdf_document
date: "2025-04-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DailyCheck #9
Slide 9: The Randomized Complete Block design
Where is the Replication occurring in this experiment?
Multiple people trying the treatments.

Where is the BLOCKING occurring in this experiment?
Each person is a block.

What should be RANDOMIZED in this experiment?
Order in which each person reveives each treatment

RCBD: 
An experimental design in which every block receives every treatment exactly 1 time.

The p-value of 0.190 is for H0: Each block has the same mean.
we should ignore this p-value, we should not let it guide what our model should be.


```{r}
person1 <- c(7, 7.5, 9, 8.25)
person2 <- c(8.25, 8.5, 9.5, 8.5)
person3 <- c(6.5, 7, 8.5, 9)
person4 <- c(8, 8, 8, 8)
wines_df <- data.frame(
score = c(person1, person2, person3, person4),
treatment = rep(c("Control", "Decanter", "Aerator", "Blender"), 4),
block = c(rep("Person1",4),
rep("Person2",4),
rep("Person3",4),
rep("Person4",4))
)
```

```{r}
library(ggplot2)
ggplot(data=wines_df, aes(y=score, x=treatment)) +
geom_point()
```

```{r}
theme_update(text = element_text(size = 20))
ggplot(data=wines_df, aes(y=score, x=treatment,
color = block, group = block)) +
geom_point() + geom_line() +
scale_x_discrete(limits=c("Control", "Decanter",
"Aerator", "Blender")) +
ggtitle("Wine tasting scores with different
methods of aeration")
```
```{r cars}
model1 <- aov(yield ~ PSI + block, data = vasc)
summary(model1)
```

```{r}
model1 <- aov(score ~ ________, data=wines_df)
summary(model1)
```

```{r}
model2 <- aov(score ~ treatment, data=wines_df)
summary(model2)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
