---
title: "DailyCheck#10"
output: pdf_document
date: "2025-04-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## Question 
1) Explain the relevance of the Sum of Squares values displayed on Slide 10 as discussed in class (Slide 12 may also be helpful in your understanding of it). 

SSE = SSB + SSE(B)
5.09375 = 2.015625 + 3.078125 
On Slide 10, we know how the total variation in our experiment is split up when we use a Randomized Complete Block Design. Without any blocking, the total leftover variation is 5.094.
We are split that variation into two parts. 
SSB(Sum of Squares for Blocks) = 2.016
SSE(B) (Residual Error after Blocking) = 3.078

the block is useful because it helps us remove extra variation that comes from differences between blocks.The leftover error becomes smaller, which improves the accuracy of our F-test. A smaller error makes it easier to spot real differences between treatments. 

```{r cars}
```


```{r}
library(readxl)
vasc <- read_excel("VascGraft.xlsx")
```


```{r}
library(dplyr)
sum_stats <- vasc %>%
group_by(PSI) %>%
summarize(mean=mean(yield), var=var(yield))
knitr::kable(sum_stats)
```
Base case:
```{r}
groupmeans <- sum_stats$mean
power.anova.test(groups = length(groupmeans), between.var = var(groupmeans), within.var = 8, 
                 power = 0.8, sig.level = 0.05, n = NULL)
```


## Including Plots
You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
