---
title: "DailyCheck#3"
output:
  pdf_document: default
  html_document: default
date: "2025-04-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
#The test statistic for a two-sample t-Test:
$$
t_0 = \frac{\bar{x}_1 - \bar{x}_2}{\sqrt{\frac{s_1^2}{n_1} + \frac{s_2^2}{n_2}}}
$$

```{r cars}

coffee <- c(8.5, 7.0, 7.25, 10.5, 6.0)
control <- c(9.5, 10.25,7.5,8.5, 8.75)
mean(coffee)
mean(control)
t.test(coffee, control)
```
```{r}
coffee_n <- length(coffee)
control_n <- length(control)

coffee_mean <- mean(coffee)
control_mean <- mean(control)

coffee_sd <- sd(coffee)
control_sd <- sd(control)

#Create summary statistics matrix
summary_stats <- matrix(c(coffee_n, coffee_mean, coffee_sd, control_n, control_mean, control_sd),
                        nrow = 2, byrow = TRUE)

#Set column and row names
colnames(summary_stats) <- c("n", "mean", "sd")
rownames(summary_stats) <- c("coffee", "control")

library(knitr)
kable(summary_stats, caption = "Summary Statistics")
```
```{r}
#Perform a t-test
model1 <- t.test(coffee, control)
names(model1)
```
```{r}
#Create a matrix for t-test result
output <- matrix(c(model1$statistic, model1$parameter, model1$p.value), nrow = 1)
colnames(output) <- c("t", "df", "p-value")
#Display the t-test results
kable(output, caption = "t-Test results")
```

```{r}
lower_bound <- round(model1$conf.int[1], 2)
upper_bound <- round(model1$conf.int[2], 2)
conf_int_sentence <- paste("The 95% confidence interval for the difference in mean sleep hours between the two groups is (", lower_bound, "," , upper_bound, ").")
conf_int_sentence
```
#The 95% confidence interval is (-3.21, 1.11). 
#This interval suggests that there is no statistically significant difference between the two groups since the interval includes zero.

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
